---
title: 'Reproducible Research: Peer Assessment 1'
author: "Sue"
date: "November 13, 2016"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set working directory and loading the necessary packages
```{r working directory, include=TRUE}
setwd('C:/Users/user/Desktop/Reproducible Research/')
library(ggplot2)
library(knitr)
library(lattice)
```

## Clearing the working environment before commencing the project
```{clear the screen, include=TRUE}
rm(list=ls(all=TRUE))
```

## Loading the data
```{r loading data,include=TRUE}
if(!file.exists('activity.csv')){unzip('activity.zip')}
activitydata=read.csv("activity.csv",stringsAsFactors=FALSE)
```
## Study the data
```{data prep, include=TRUE}
head(activitydata)
str(activitydata)
```

## Mean and median total number of steps taken per day
```{mean and median, include=TRUE}
steps_by_day<-aggregate(steps~date,activitydata,sum)
rmean<-mean(steps_by_day$steps)
rmedian<-median(steps_by_day$steps)
```
The mean is 10766.19
The median is 10765

## Including Plots
```{r plot, echo=TRUE}
steps_by_interval<-aggregate(steps~interval,activitydata,mean)

plot( x=steps_by_interval[,1], 
      y=steps_by_interval[,2], 
      type="l",
      col="black",
      main="Average steps recorded per interval",
      ylab="Amount of steps", 
      xlab="Time interval")

max_interval<-steps_by_interval[which.max(steps_by_interval$steps),1]
```
The 5-minute interval that, on average, contains the maximum number of steps is 835

## Imputing missing values and data cleaning
```{r plot 2, echo=TRUE}

incomplete<-sum(!complete.cases(activitydata))

Newdata<-transform(activitydata,steps=ifelse(is.na(activitydata$steps),steps_by_interval$steps[match(activitydata$interval,steps_by_interval$interval)],activitydata$steps))

Newdata[as.character(Newdata$date)=="22-10-01",1]<-0

steps_by_day2<-aggregate(steps~date,Newdata,sum)
hist(steps_by_day2$steps,main=paste("Total steps taken each day"),col="green",xlab="Number of Steps")
```

##Are there differences in activity patterns between weekdays and weekends
```{r plot 3, echo=TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday","Thursday", "Friday")

Newdata$dateofweek=as.factor(ifelse(is.element(weekdays(as.Date(Newdata$date)),weekdays),"Weekday","Weekend"))

steps_by_interval_i<-aggregate(steps ~ interval + dateofweek, Newdata, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dateofweek, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```
